Speakers of diverse languages structure their utterances for efficient communication

### 2019-09-24 Paper/info-curves.Rmd
first couple of paras

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Speakers of diverse languages structure their utterances for efficient communication},
            pdfauthor={Josef Klafka and Daniel Yurovsky},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Speakers of diverse languages structure their utterances for efficient communication}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Josef Klafka and Daniel Yurovsky}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{2019-09-17}

\usepackage{setspace}

\doublespacing

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

What role does communicative efficiency play in how we organize our utterances? In this paper, we present a novel method of examining how much information speakers in a given language communicate in each word in their utterances, surveying numerous diverse languages. We find that speakers produce frequent and informative words at regular parts of their utterances, depending on language they use, which is predictable in part from the features and genealogy of their language. This robust information distribution characterizes both spoken and written communication, and emerges in children's earliest utterances. However, in real-time communication, in-context word predictability allows listeners to process information at a constant, optimal rate, regardless of the information distribution in the language they understand.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

We use language to communicate information. Although we often need to say things we have never said before, we are easily and quickly able to produce novel utterances with next to no prior planning. Our listeners incrementally process each word we say in utterances they may have never heard before, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g. Kutas \& Federmeier (2011)). Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication. Listeners may struggle to decode unpredictable words, which can cause difficulty decoding the following words as well. Efficient and effective communication relies on interlocutors being able to process novel information without being surprised by new information in unpredictable words. On the other hand, to make best use of the capacity of the communication channel, speakers should encode as much information as possible into each word they say.

Due to their differences in distributional frequency, words vary in how much information content they contain (Shannon, 1948). Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be (Zipf, 1949). Over and above frequency, contextual information content predicts word length, due to speakers minimizing effort in the amount of information they encode in speech (Piantadosi, Tily, \& Gibson, 2011). Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side (Jaeger \& Levy, 2007). Across utterances, context allows speakers to produce more informative and surprising sentences later in communication. The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains (Genzel \& Charniak, 2002). From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.

The pressure for efficient communication influences language at all timescales, from online speech production to evolution. Mahowald, Fedorenko, Piantadosi, \& Gibson (2013) found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility. Predictability and ambiguity shape the distribution of color terms in languages (Gibson et al., 2017) as well as word lengths (Piantadosi et al., 2011) across languages, and speakers are able to deterministically decode words despite ambiguity in word forms (Piantadosi, Tily, \& Gibson, 2012).

How do frequency and predictive context determine how we distribution information at the level of utterances? We do not speak only in phrases or paragraphs, but conversationally we speak in utterances and in writing we produce sentences as a coherent unit of a complete thought. Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages. The average rate of information speakers transfer in syllables differs from language to language (Aylett \& Turk, 2004). How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?

Yu, Cong, Liang, \& Liu (2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances. Treating each word position by itself, Yu et al. (2016) calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty (Shannon, 1948). They found a distinctive three-step shape to the distribution of information in English written sentences.

In this paper, we expand on this body of prior work in a number of novel ways. We replicate the results from Yu et al. (2016) with a different metric, one intimately tied to incremental word processing. We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech. We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words. We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information. We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax. Speakers of a language will tend to distribute information according to that language's distribution of information in their utterances. As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Shannon (1948) defined information as ``the reduction in uncertainty about one variable given that the value of another variable is known''. We use a metric proposed by Shannon and applied by Levy (2008): lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen. Equivalently, we can compute surprisal with the predictability of the word, as in the formula below. The surprisal of a word is inversely proportional to the predictability of a word, such that less common and less predictable words carry more information.

\[surprisal(word) = -log P(word)\]

The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking (Smith \& Levy, 2013) and ERP (Frank, Otten, Galli, \& Vigliocco, 2015) studies, among other sources. Considered without context, the surprisal of an individual word is inversely proportional to the frequency of that word, so that the less often a person has seen a word, the more information that word holds. For example, ``flower'' has less information than ``azalea'' because ``flower'' is much more common than ``azalea''. Though the two words have the same length in number of letters, it's more difficult to process ``azalea'' when reading it here than when reading ``flower''. Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically (Piantadosi et al., 2011).

However, when reading or listening, people don't just consider each word as an isolated linguistic signal. Instead, listeners use the words they have already heard to predict and decode the word they are currently processing. Following this incremental processing paradigm, we can also condition the surprisal of a word in its context. In the formula below, widenotes the word currently being read or heard, while wi-1denotes the first word before the current word, wi-2denotes the second word before the current word, and so on.

\[surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)\]
\[= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}\]

When we use a word or two of context in our surprisal calculations, then the set of reasonable final items in our ngrams is greatly restricted. ``Flower'' may contain less information than ``azalea'' when we consider the words independently of their context, but with context this can be reversed. Flower appears in a variety of contexts, and so the information content of a word like ``flower'' in a particular context may be higher than ``azalea''. If you only have azaleas in your garden, then hearing someone say ``in that garden, look at the flowers'' may be higher surprisal for you: you expect them to say ``azalea''. This prediction does not need many words before to work out, for example in ``I take my coffee with cream and sugar''. When you hear ``cream and'', you automatically predict ``sugar'' or maybe ``honey'', but there are few possible continuations with even those two words. Hearing ``I'' restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.

@@

The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e.~independently information-rich words. The context-based surprisal metric show us how speakers tend to distribute the information in utterances relative to real-time processing in communication. We expect a priori that our frequency-based surprisal curve will be flat. No one part of the sentence will on average have words that are more frequent than another across utterance lengths. Similarly, we expect that there will be a small smoothing effect for our contextual surprisal metric.

\hypertarget{written-and-spoken-english-have-the-same-information-distribution}{%
\section{Written and Spoken English have the same information distribution}\label{written-and-spoken-english-have-the-same-information-distribution}}

We first turn to working with written English in the British National Corpus (BNC; Leech, 1992). The BNC is a collection of spoken and written records (90\% written) from the turn of the century, intended to be a representative sample of British English. Using their word entropy metric without context, Yu et al.~found a distinctive three-step distribution for information in written English sentences in the corpus. The first word tended to contain little information. While the middle words of sentences each had more information than the first word, information was flat and non-increasing across the middle of sentences. The final word, they found, contained the most information out of any in the sentence, with a noticeable spike in information. They found the same distribution across sentence lengths, from sentences with 15 words to sentences with 45 words.

We first replicate the Yu et al. (2016) result using the surprisal metric in place of the entropy metric. We use the frequency-based or ``contextless'' surprisal metric, which determines the average distribution of information based on word frequencies in a corpus. A priori we expect that the frequency-based metric will produce a flat distribution of information across word positions in the BNC.

See in Figure @ref(fig:bnc\_unigrams) a qualitative comparison of the information trajectory they found and the information trajectory we found. It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al. (2016) did. The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/bnc_unigrams-1.pdf}
\caption{(\#fig:bnc\_unigrams)BNC frequency-based information curves}
\end{figure}

@@

For Switchboard we find the same information trajectory as we did for the BNC, and we find the same distribution of information regardless of utterance length. The longer the utterance, the more pronounced the information distribution shape becomes. A qualitative comparison of the two trajectories is in Figure @ref(fig:switchboard\_unigrams).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/switchboard_unigrams-1.pdf}
\caption{(\#fig:switchboard\_unigrams)Switchboard frequency-based information curves}
\end{figure}

We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length. This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information. The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from (Genzel \& Charniak, 2002).

What about context? So far we've only discussed the frequency-based metric, considering words on their own without any explicit incorporation of prior context. As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process. We now include one word of context (bigrams) and then two words of context (trigrams) for each word in our measurements. We observe a flattening effect of context across both modalities and all speaker populations. After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate. The contextual information curves for the BNC and Switchboard are in Figures @ref(fig:switchboard\_bigrams) and @ref(fig:switchboard\_trigrams).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/switchboard_bigrams-1.pdf}
\caption{(\#fig:switchboard\_bigrams)Switchboard bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/switchboard_trigrams-1.pdf}
\caption{(\#fig:switchboard\_trigrams)Switchboard trigram context-based information curves}
\end{figure}

@@

\hypertarget{english-cds-and-child-speech}{%
\section{English CDS and child speech}\label{english-cds-and-child-speech}}

@@

We use the Providence corpus from CHILDES (Evans \& Demuth, 2012; MacWhinney, 2000), which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents. We obtained the Providence corpus using the childesr frontend to the childes-db database (Sanchez et al., 2019). The child age range was between 11 months and 4 years. Sessions recorded in the home. Parents said a majority of the utterances, but children take up an increasing share of the conversation as they grow older and are actually able to speak full utterances. The utterances in the Providence corpus are on average significantly shorter than those in the BNC and Switchboard; over 90\% of the utterances in the Providence corpus are at most 10 words long, with the median at around 6 words. Similar to our treatment of the Switchboard corpus, we split the Providence corpus up by conversational turns and separate utterances.

We observe the same distinctive distribution of information for parents and children in the Providence corpus as we did for adults in the BNC and Switchboard. The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances. See @ref(fig:providence\_unigrams).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/providence_unigrams-1.pdf}
\caption{(\#fig:providence\_unigrams)Providence frequency-based information curves. Lines around each point indicate 95\% confidence intervals computed with non-parametric bootstrap}
\end{figure}

What about context in the Providence corpus? When incorporating one or two words of predictive context, we observe the same trend as in the Switchboard and BNC corpora. Beyond the first couple of words, once your interlocutor has enough context to predict with some accuracy what you will say next, then you decode information from their speech stream at a constant and optimal rate. This applies to parents and children speaking to one another, as well as adults speaking and writing to one another. See @ref(fig:providence\_bigrams) and @ref(fig:providence\_trigrams).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/providence_bigrams-1.pdf}
\caption{(\#fig:providence\_bigrams)Providence bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/providence_trigrams-1.pdf}
\caption{(\#fig:providence\_trigrams)Providence trigram context-based information curves}
\end{figure}

\hypertarget{child-speech-and-cds-across-languages}{%
\section{Child speech and CDS across languages}\label{child-speech-and-cds-across-languages}}

>>>>

---
title             : "Speakers of diverse languages structure their utterances for efficient communication"
shorttitle        : "Cross-linguistic information structure"

author: 
  - name          : "Josef Klafka"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "5000 Forbes Ave"
    email         : "klafka@cmu.edu"
  - name          : "Daniel Yurovksy"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Carnegie Mellon University"
  - id            : "2"
    institution   : "University of Chicago"

authornote: |
  Thanks to folks. JSMF.

abstract: |
  What role does communicative efficiency play in how we organize our utterances? In this paper, we present a novel method of examining how much information speakers in a given language communicate in each word in their utterances, surveying numerous diverse languages. We find that speakers produce frequent and informative words at regular parts of their utterances, depending on language they use, which is predictable in part from the features and genealogy of their language. This robust information distribution characterizes both spoken and written communication, and emerges in children's earliest utterances. However, in real-time communication, in-context word predictability allows listeners to process information at a constant, optimal rate, regardless of the information distribution in the language they understand.

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["info-curves.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library(xtable)
library(knitr)
library(papaja)
library(ggthemes)
library(lmerTest)
library(directlabels)
library(ggrepel)
library(feather)
library(here)
library(tidyboot)
library(broom)
library(broom.mixed)
library(ggdendro)
library(gridExtra)
library(ggpmisc)

library(tidyverse)

theme_set(theme_classic(base_size = 14) + theme(legend.position = "none"))

knitr::opts_chunk$set(fig.pos = 'tb', echo = FALSE, warning = FALSE, 
                      cache = TRUE, warning = FALSE, message = FALSE, 
                      sanitize = TRUE, fig.path='figs/', fig.width = 3,
                      fig.height = 3)
set.seed(42)
options(digits=2)
```

```{r load data, include = F}
childes_results <- read_feather(here("Data/all_childes_bootstrap.feather"))
wiki_results <- read_feather(here("../surprisal/Data/all_spaces.feather"))
bnc_results <- read_feather(here("Data/bnc_bootstrap.feather"))
switch_results <- read_feather(here("Data/switch_bootstrap.feather"))
wiki_curves <- read_csv(here("../surprisal/Data/relative_ngrams.csv"))
wiki_lr <- read_feather(here("../surprisal/Wikipedia/models.feather"))
```

# Introduction

One of the defining features of human language is its power to transmit information. We use language for a variety of different proximal tasks tasks: greeting friends, taking notes, signalling group identities. But all share a common unifying purpose: changing the mental state of the listener or reader of our communications [@austin1975]. Language can thus naturally thought of as a code that allows speakers to turn their intended meanings into a form that can be transmitted through the air (or paper, or electrons) and subsequently converted by listeners back into an approximation of the intended meaning [@shannon1948].

Beyond its utility as a metaphor, this coding perspective on language is powerful because it affords a framework for rational analysis: If language has evolved to be an optimal code for information transmission, how should this code be structured [@anderson1989]? The optimal code would have to optimize two competing pressures: (1) a pressure for listeners to easily and successfully decode messages sent by the speaker, and (2) a pressure for speakers to easily code their messages and transmit them with minimal effort and error. A fundamental constriaint on both of these processes is the serial nature of human language: sounds are produced one at a time and each is unavailable perceptually once it is no longer being produced.

The solution that languages have arrived on is incremental processing: Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion [@kutas2011; @tanenhaus1995; @pickering2013]. Because prediction errors can lead to severe processing costs and difficulty integrating new information, an optimal coding strategy is for speakers seeking to minimize their prodction costs is to produce utterances that are are just at the prediction capacity of listeners without exceeding this capacity [@aylett2004; @genzel2002].

One salient constraint on this transmission processes is the serial nature of language:

This coding perspective is powerful because it lends itself naturally to a rat

Languages can thus be thought of as a code that can be used to transform speaker's intented meaning into a

* Coding/decoding and information transfer
* language as prediction
* how should the code be structured? optimization questions like this have a had * a lot power across different parts of language: zipf, kemp & regier, etc etc.  noisy channel stuff. (or does this go in discussion)
* don't under or overutilize channel, stay near limit -- constant entropy rate
* evidence for this at multiple levels from individiual lexical choice *(info/information theory paper), inclusion of non-obligatory aspects of language (filled pauses, that), across multiple utterances in a discourse
* what about over the course of an individual utterance--perhaps the strongest constraints from language itself?
* yu et al. found this puzzling effect. 
* we replicate that in a different kind of analysis and show that it extends to trigrams etc???
* BUT variation cross-linguistically. why? constraints from language.

We use language to communicate information. Although we often need to say things we have never said before, we are easily and quickly able to produce novel utterances with next to no prior planning. Our listeners incrementally process each word we say in utterances they may have never heard before, and predict which words will come next within fractions of a second of hearing the beginning of the word [for evidence see e.g. @kutas2011]. Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication. Listeners may struggle to decode unpredictable words, which can cause difficulty decoding the following words as well. Efficient and effective communication relies on interlocutors being able to process novel information without being surprised by new information in unpredictable words. On the other hand, to make best use of the capacity of the communication channel, speakers should encode as much information as possible into each word they say.

Due to their differences in distributional frequency, words vary in how much information content they contain [@shannon1948]. Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be [@zipf1949]. Over and above frequency, contextual information content predicts word length, due to speakers minimizing effort in the amount of information they encode in speech [@piantadosi2011]. Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side [@jaeger2007]. Across utterances, context allows speakers to produce more informative and surprising sentences later in communication. The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains [@genzel2002]. From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.

The pressure for efficient communication influences language at all timescales, from online speech production to evolution. @mahowald2013 found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility. Predictability and ambiguity shape the distribution of color terms in languages [@gibson2017] as well as word lengths [@piantadosi2011] across languages, and speakers are able to deterministically decode words despite ambiguity in word forms [@piantadosi2012].

How do frequency and predictive context determine how we distribution information at the level of utterances? We do not speak only in phrases or paragraphs, but conversationally we speak in utterances and in writing we produce sentences as a coherent unit of a complete thought. Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages. The average rate of information speakers transfer in syllables differs from language to language [@aylett2004]. How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?

@yu2016 proposed a novel method for quantifying the amount of information typically found at each part of English utterances. Treating each word position by itself, @yu2016 calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty [@shannon1948]. They found a distinctive three-step shape to the distribution of information in English written sentences.

In this paper, we expand on this body of prior work in a number of novel ways. We replicate the results from @yu2016 with a different metric, one intimately tied to incremental word processing. We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech. We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words. We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information. We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax. Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances. As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.

# Methods

@shannon1948 defined information as "the reduction in uncertainty about one variable given that the value of another variable is known". We use a metric proposed by Shannon and applied by @levy2008: lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen. Equivalently, we can compute surprisal with the predictability of the word, as in the formula below. The surprisal of a word is inversely proportional to the predictability of a word, such that less common and less predictable words carry more information.

$$surprisal(word) = -log P(word)$$

The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking [@smith2013] and ERP [@frank2015] studies, among other sources. Considered without context, the surprisal of an individual word is inversely proportional to the frequency of that word, so that the less often a person has seen a word, the more information that word holds. For example, “flower” has less information than “azalea” because “flower” is much more common than “azalea”. Though the two words have the same length in number of letters, it’s more difficult to process “azalea” when reading it here than when reading “flower”. Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically [@piantadosi2011].

However, when reading or listening, people don’t just consider each word as an isolated linguistic signal. Instead, listeners use the words they have already heard to predict and decode the word they are currently processing. Following this incremental processing paradigm, we can also condition the surprisal of a word in its context. In the formula below, widenotes the word currently being read or heard, while wi-1denotes the first word before the current word, wi-2denotes the second word before the current word, and so on.

$$surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)$$
$$= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}$$

When we use a word or two of context in our surprisal calculations, then the set of reasonable final items in our ngrams is greatly restricted. “Flower” may contain less information than “azalea” when we consider the words independently of their context, but with context this can be reversed. Flower appears in a variety of contexts, and so the information content of a word like “flower” in a particular context may be higher than “azalea”. If you only have azaleas in your garden, then hearing someone say “in that garden, look at the flowers” may be higher surprisal for you: you expect them to say “azalea”. This prediction does not need many words before to work out, for example in “I take my coffee with cream and sugar”. When you hear “cream and”, you automatically predict “sugar” or maybe “honey”, but there are few possible continuations with even those two words. Hearing “I” restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.

@@

The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e. independently information-rich words. The context-based surprisal metric show us how speakers tend to distribute the information in utterances relative to real-time processing in communication. We expect a priori that our frequency-based surprisal curve will be flat. No one part of the sentence will on average have words that are more frequent than another across utterance lengths. Similarly, we expect that there will be a small smoothing effect for our contextual surprisal metric.

# Written and Spoken English have the same information distribution

We first turn to working with written English in the British National Corpus [BNC; @leech1992]. The BNC is a collection of spoken and written records (90% written) from the turn of the century, intended to be a representative sample of British English. Using their word entropy metric without context, Yu et al. found a distinctive three-step distribution for information in written English sentences in the corpus. The first word tended to contain little information. While the middle words of sentences each had more information than the first word, information was flat and non-increasing across the middle of sentences. The final word, they found, contained the most information out of any in the sentence, with a noticeable spike in information. They found the same distribution across sentence lengths, from sentences with 15 words to sentences with 45 words.

We first replicate the @yu2016 result using the surprisal metric in place of the entropy metric. We use the frequency-based or “contextless” surprisal metric, which determines the average distribution of information based on word frequencies in a corpus. A priori we expect that the frequency-based metric will produce a flat distribution of information across word positions in the BNC.

See in Figure \@ref(fig:bnc_unigrams) a qualitative comparison of the information trajectory they found and the information trajectory we found. It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as @yu2016 did. The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.

```{r bnc_unigrams, echo = F, fig.height=2, fig.cap="BNC frequency-based information curves"}
bnc_results %>% 
  filter(context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, 
             ymin = ci_lower, ymax = ci_upper, color = as.factor(length))) + 
    geom_pointrange() + 
    geom_line() +
    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

For Switchboard we find the same information trajectory as we did for the BNC, and we find the same distribution of information regardless of utterance length. The longer the utterance, the more pronounced the information distribution shape becomes. A qualitative comparison of the two trajectories is in Figure \@ref(fig:switchboard_unigrams).

```{r switchboard_unigrams, echo=F, fig.height=4, fig.cap="Switchboard frequency-based information curves"}
sw_plot <- switch_results %>% 
  filter(context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")

bnc_plot <- bnc_results %>% 
  filter(context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
    xlab("Word position") + 
    ylab("Mean information")

grid.arrange(bnc_plot, sw_plot, nrow = 2)
```

We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length. This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information. The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from [@genzel2002].

What about context? So far we’ve only discussed the frequency-based metric, considering words on their own without any explicit incorporation of prior context. As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process. We now include one word of context (bigrams) and then two words of context (trigrams) for each word in our measurements. We observe a flattening effect of context across both modalities and all speaker populations. After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate. The contextual information curves for the BNC and Switchboard are in Figures \@ref(fig:switchboard_bigrams) and \@ref(fig:switchboard_trigrams).

```{r switchboard_bigrams, echo=F,fig.height=4,fig.cap="Switchboard bigram context-based information curves"}
sw_plot <- switch_results %>% 
  filter(context == 1) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")

bnc_plot <- bnc_results %>% 
  filter(context == 1) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
    xlab("Word position") + 
    ylab("Mean information")

grid.arrange(bnc_plot, sw_plot, nrow = 2)
```

```{r switchboard_trigrams, echo=F,fig.height=4,fig.cap="Switchboard trigram context-based information curves"}
sw_plot <- switch_results %>% 
  filter(context == 2) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")

bnc_plot <- bnc_results %>% 
  filter(context == 2) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
    xlab("Word position") + 
    ylab("Mean information")

grid.arrange(bnc_plot, sw_plot, nrow = 2)
```

@@

# English CDS and child speech

@@

We use the Providence corpus from CHILDES [@macwhinney2000; @evans2012], which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents. We obtained the Providence corpus using the childesr frontend to the childes-db database [@sanchez2019]. The child age range was between 11 months and 4 years. Sessions recorded in the home. Parents said a majority of the utterances, but children take up an increasing share of the conversation as they grow older and are actually able to speak full utterances. The utterances in the Providence corpus are on average significantly shorter than those in the BNC and Switchboard; over 90% of the utterances in the Providence corpus are at most 10 words long, with the median at around 6 words. Similar to our treatment of the Switchboard corpus, we split the Providence corpus up by conversational turns and separate utterances.

We observe the same distinctive distribution of information for parents and children in the Providence corpus as we did for adults in the BNC and Switchboard. The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances. See \@ref(fig:providence_unigrams).

```{r providence_unigrams, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves. Lines around each point indicate 95% confidence intervals computed with non-parametric bootstrap"}
childes_results %>% 
  filter(language == "English", context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, 
             ymax = ci_upper, 
             color = speaker)) + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

What about context in the Providence corpus? When incorporating one or two words of predictive context, we observe the same trend as in the Switchboard and BNC corpora. Beyond the first couple of words, once your interlocutor has enough context to predict with some accuracy what you will say next, then you decode information from their speech stream at a constant and optimal rate. This applies to parents and children speaking to one another, as well as adults speaking and writing to one another. See \@ref(fig:providence_bigrams) and \@ref(fig:providence_trigrams).

```{r providence_bigrams, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
childes_results %>% 
  filter(language == "English", context == 1) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```
